\documentclass{article}

% need to procedurally generate this file...

% this is just from my preamble; 
% delete since this won't exist for you
\input{~/docs/educ/latex/class/packages.tex}
\input{~/docs/educ/latex/class/style.tex}
\input{~/docs/educ/latex/class/section_styles.tex}
\input{~/docs/educ/latex/class/fixes.tex}
% delete all of the above \input stuff

\title{Math 336 Homework 1}

% reals
 \newcommand\R{\ensuremath{\mathbb{R}}}
% complex
 \newcommand\C{\ensuremath{\mathbb{C}}}
% disc
 \newcommand\D{\ensuremath{\mathbb{D}}}

% Re & Im operators
\newcommand{\re}{\operatorname{Re}}
\newcommand{\im}{\operatorname{Im}}

\begin{document}
\maketitle
\begin{problem}
Describe geometrically the sets of points $z$ in the complex plane defined by the
following relations:
\begin{enumerate}[label=(\alph*)]
	\item $|z-z_1| = |z-z_2|$ where $z_1,z_2\in \C$.
	\item $1 / z = \overline{z}$.
	\item $\re(z)=3$.
	\item $\re(z)>c$, (resp., $\ge c$) where $c\in \R$.
	\item $\re(az+b)>0$ where $a,b \in \C$.
	\item $|z| =\re(z)+1$.
	\item $\im(z)=c$ with $c \in \R$.
\end{enumerate}
\end{problem}
\begin{problem}
With $\omega = se^{i\varphi}$, where $s\ge 0$ and $\varphi \in \R$, solve the equation $z^n = \omega$ in $\C$ where $n$ is a natural number.
How many solutions are there?
\end{problem}
\begin{problem}
The family of mappings introduced here plays an important role in complex
analysis.
These mappings, sometimes called \textbf{Blaschke factors}, will reappear in
various applications in later chapters.
\begin{enumerate}[label = (\alph*)]
	\item Let $z,w$ be two complex numbers such that  $\overline{z} w \neq 1$.
	      Prove that \[
		      \left| \frac{w-z}{1-\overline{w}z} \right| <1\quad \text{if } |z|<1 \text{ and } |w| < 1
		      ,\] and also that \[
		      \left| \frac{w-z}{1-\overline{w}z} \right| =1\quad \text{if } |z|=1 \text{ or } |w| = 1
		      .\]
	      [Hint: Why can one assume that $z$ is real?
	      It then suffices to prove that  \[
		      (r-w)(r-\overline{w})\le (1-rw)(1-r\overline{w})
	      \] with equality appropriate for $r$ and  $|w|$.]
	\item Prove that for a fixed $w$ in the unit disc  $\D$, the mapping \[
		      F: z \mapsto \frac{w-z}{1-\overline{w}z}
	      \] satisfies the following conditions:
	      \begin{enumerate}[label = (\roman*)]
		      \item $F$ maps the unit disc to itself (that is, $F:\D \to \D$), and is holomorphic.
		      \item $F$ interchanges 0 and $w$, namely $F(0)=w$ and $F(w)=0$.
		      \item $|F(z)| = 1$ if $|z| = 1$.
		      \item $F:\D \to \D$ is bijective.
			            [Hint: Calculate $F \circ F$]
	      \end{enumerate}
\end{enumerate}
\end{problem}
\begin{problem}
Consider the function defined by \[
	f(x+iy)=\sqrt{|x||y|},\quad \text{whenever } x,y \in \R
	.\]
Show that $f$ satisfies the Cauchy-Reimann equations at the origin, yet $f$ is not holomorphic at 0.
\end{problem}
\begin{problem}
In this problem we will go through a proof of the Fundamental Theorem of Algebra,
that is: If \[
	p(z)=a_n z^n + \cdots + a_0
\] is a polynomial with an $a_n\neq 0$, then there exists $z_0\in \C$ such that $p(z_0)=0$.
\begin{enumerate}[label=(\roman*)]
	\item Suppose for the sake of contradiction that $p(z) \neq 0$ for all $z \in \C$. Show that the function $g(z) = |p(z)|$ has a minimum at some point $z_0\in \C$. (Hint: Remember that $\C$ is definitely not compact!)
	\item Consider the function $q(z)=\frac{1}{|p(z_0)|}p(z+z_0)$.
	      Show that $q$ is a polynomial with $|q(0)| = 1$ and that $|q(z)|$ has its minimum at $z = 0$.
	\item Show that for any sufficiently small $\epsilon>0$, there is some $\theta$ for which $|q(\epsilon e^{i\theta})|<1$, which provides the desired contradiction.
\end{enumerate}
\end{problem}
\begin{problem}
Consider the function $f$ defined on $R$ by \[
	f(x) = \begin{cases}
		0            & \text{if } x\le 0, \\
		e^{-1 / x^2} & \text{if } x > 0.
	\end{cases}
\]
Prove that $f$ is indefinitely differentiable on $R$, and that $f^{(n)}(0)=0$ for all $n\ge 1$.
Conclude that $f$ does not have a converging power series expansion $\sum_{n=0}^{\infty} a_n x^n$ for $x$ near the origin.
\end{problem}
\begin{problem}
Show that if $|a|<r|b|$, then  \[
	\int_{\gamma}\frac{1}{(z-a)(z-b)}dz = \frac{2\pi i}{a-b}
	,\] where $\gamma$ denotes the circle centered at the origin, of radius $r$, with the positive orientation.
% note that the previous parts of this question in the textbook have that, for all integers n (except n=-1), the integral about a circular path anywhere in the complex plane is zero; at n=-1, the integral of a circular path about the origin is 2 pi i. 
\end{problem}
\end{document}
