\documentclass[../9extra]{subfiles}
\begin{document}
\begin{problem}[1]
Let $K\subset \R^{n}$ be compact.
Let $f:K\longrightarrow K$ be a \textit{shrinking map}.
$\forall x,y \in K,\, x \neq y \implies |f(x)-f(y)| < |x-y|$.

Prove that $f$ has a unique fixed point  $x \in K : x = f(x)$.
\end{problem}

If $K$ compact with  $K \supset C_1 \supset C_2 \supset \cdots$ nested sequence of non-empty closed subsets $C_i$, then \[
	\bigcap_{i=1}^{\infty} C_i \neq \O \tag{*}
	.\]

\begin{proposition}[Continuity of the shrinking map]
	$f$ is uniformly continuous.
\end{proposition}
\begin{proof}[Proof of Continuity Proposition]
	Choose $\delta = \epsilon$.
	Then, $\forall \epsilon>0,\, \exists \delta>0$, \[
		|f(x)-f(y)| < |x-y| < \delta = \epsilon
		.\]
	So, $f$ is uniformly continuous as  $\delta$ depends solely on  $\epsilon$.
\end{proof}

\begin{proposition}[Uniqueness of the fixed point]
	The fixed point of $f$ is unique.
\end{proposition}
\begin{proof}[Proof of Uniqueness Proposition]
	Suppose, for a contradiction, that there were $x,y$ fixed points with  $x \neq y$
	and $f(x)=x$,  $f(y)=y$.

	Then,  $|f(x)-f(y)| = |x-y|$.
	But we had, by the definition of  $f$, $|f(x)-f(y)| < |x-y|$, which is a contradiction.

	Thus, $x=y$ and the fixed point is unique.
\end{proof}

\begin{proof}[Proof of Problem]
	Let $g : K \longrightarrow \R$ be the distance between $x$ and its image under $f$, \[
		g(x)= |f(x)-x|
		.\]
	Since $f$ is continuous, then $g$, the difference between two continuous functions, is also continuous.

	Since  $K$ is compact and  $g$ is continuous and positive, then, by EVT, there is a minimum of $g$ at $x \in K$, \[
		0\le \underset{x \in K}{\min{g}} = g(a)
		.\]
	We wish to show that $g(a)$ must be zero.
	This $a$ is the fixed point of $f$.

	First, if $g(a)=0$, then we're done.

	Next, if  $g(a)>0$, we will consider the image of $f(a)$ under $g$.
	But, the definition of $f$ bounds this value, \[
		g(f(a))=|f(a)-f(f(a))| < |a - f(a)| = g(a)
		.\]

	So, we have that $g(f(a))<g(a)$, contracting the fact that $g(a)$ was the minimum of  $g$.

	Thus,  $g(a)=0$ and the fixed point of  $f$ occurs at  $x=a$.
\end{proof}
\end{document}
