\documentclass[../hw5]{subfiles}
\begin{document}
\begin{problem}
Determine whether the graphs from Problems 2 and 3 of Homework 4 are bipartite.

Let $d_H(x,y)$ be the Hamming distance between two binary strings, which represents the number of bit positions in which  $x$ and  $y$ differ.

We can label each binary string with an even or odd parity based on the sum of bits mod 2 (0 is even parity and 1 is odd parity).
\begin{enumerate}[label=\alph*)]
	\item $G_1=\{\{0,1\}^n, \{xy\ \mid\ x,y \in V,\ d_H(x,y)=1 \}  \} $.
	\item $G_2=\{\{0,1\}^n, \{xy\ \mid\ x,y \in V,\ d_H(x,y)=2 \}  \} $.
\end{enumerate}
\end{problem}
\begin{proof}[Proof of a]
	We claim that $G$ is bipartite by vertex parity.

	Since every edge connects vertices of opposite parity, we must have a bipartition $\{X,Y\}$ where $X$ contains vertices of even parity and $Y$ contains vertices of odd parity.

	Since there are no edges between vertices of the same parity, we have established a bipartition.
\end{proof}
\begin{proof}[Proof of b]
	We will first consider $n=1$.
	0 and 1 are the only vertices.
	Since they do not differ by two bits, they have no edge between them and are thus bipartite.

	For  $n=2$, we have that 01 and 10 are connected, as well as 00 and 11.
	So, we have a bipartition  $\{\{01,00\},\, \{10,11\}\}$ of $G_2$ when $n=2$.

	For $n\ge 3$, we have a three cycle or vertices that contain the following substrings: 001, 010, and 100, each of which differ by two bits, so they form an odd cycle.
	This set of substrings has the same parity, they must be contained in at least one of the connected components of $G_2$ when $n\ge 3$.

	Since there is an odd cycle in $G_2$, then $G_2$ is not bipartite for $n\ge 3$.
\end{proof}
\end{document}
