\documentclass[../astro_4]{subfiles}
\begin{document}
\begin{problem}
Which gases from Table 3 could Mars retain?
\end{problem}
Mars has an escape velocity of 5000 m/s, so it can contain gases moving at less than $\frac{1}{6}5000 \approx 833$ m/s.

Mars is located 1.5 AU away from the Sun, this temperature is not directly given on Table 2,
but we expect it to be between 283 and 400 K (the temps for 2 and 1 AU respectively).

We can research the average temperature on mars to see that is is slightly colder, at around 250 K.
\footnote{\url{https://www.britannica.com/question/What-is-the-temperature-on-Mars}.}

However, we used a value of 400 K for the temperature of earth for this exercise, that's equivalent to around 260 F, which is much hotter than the actual earth.

So, we should attempt to determine what temperature Table 2 would give at a distance of 1.5 AU, noting that our value ought to be within the range [283,400].

We can create an regression given the data in Table 2 of the form $T~\frac{a}{\sqrt{r}}+b$, where $T$ is temperature, $r$ is distance from the world to the sun, and $a$ and $b$ are constants.

We arrive at this form given that the energy received per area is inversely proportional to the square of distance.

But, our form has $\frac{1}{\sqrt{r}}$ and not $\frac{1}{r^2}$; we also must consider the Stefano-Boltzmann law that energy is proportional to the fourth power of temperature.

Thus, $\frac{1}{r^2}\propto T^4 \implies T \propto \frac{1}{r^{1 / 2}}$.

Our model now produces a graph with an $R^2$ of about 0.9998, coefficients $a=392.41373$ and $b=3.83379$, and a predicted temperature of 324 K at 1.5 AU.
\footnote{Using Desmos graphing calculator.}

We will compute the average velocity for each gas in Table 3 as if it were in the Martian atmosphere at 324 K.
\[
	\begin{array}{c | c}
		\text{gas}         & \text{average velocity (m/s)}           \\
		\hline
		\text{H}_2         & 157\sqrt{\frac{324}{2}} \approx 1998    \\
		\text{CH}_4        & 157\sqrt{\frac{324}{16}} \approx 707    \\
		\text{NH}_3        & 157\sqrt{\frac{324}{17}} \approx    685 \\
		\text{H}_2\text{0} & 157\sqrt{\frac{324}{18}} \approx   666  \\
		\text{N}_2         & 157\sqrt{\frac{324}{28}} \approx   534  \\
		\text{CO}_2        & 157\sqrt{\frac{324}{44}} \approx  426   \\
	\end{array}
\]
Since Mars has an escape velocity of 833 m/s, it can retain all gases within the table moving slower, which include all but hydrogen.
\end{document}
